---
title: "Iran Dental Schools Ranking"
author: "Ahmad Sofi-Mahmudi"
format: pdf
editor: visual
---

## Aim

In this project, I aimed to rank dental schools in Iran using principal component analysis of five variables:

-   Average mark of students;

-   Number of all professors;

-   Number of full and associate professors;

-   Median mark in residency exam; and,

-   The proportion of participants in the residency exam who were allowed to choose their specialty.

## Analysis

First, we load the data:

```{r}
dental_data = read.csv("iran_dental_schools_data.csv")
```

This data set has data for these 34 schools:

```{r}
dental_data$School
```

The variables are:

```{r}
names(dental_data)
```

Now, we run the PCA analysis:

```{r}
library(FactoMineR)
library(dplyr)

pca_out = PCA(dental_data[,-1])
score0 = pca_out$ind$coord[,1]
pca_score = 100*(score0 - min(score0))/(max(score0) - min(score0))
score = as.data.frame(pca_score)
pca_data = bind_cols(dental_data, score)
```

Here are the scores:

```{r}
sorted_pca = pca_data[, c(1, 7)][order(pca_data$pca_score, decreasing = T),]
sorted_pca
```

We can also plot it:

```{r}
library(ggplot2)

sorted_pca %>% 
        ggplot() +
        aes(x = reorder(School, pca_score), y = pca_score) +
        geom_bar(stat = "identity") +
        scale_fill_gradient2(high = "red", low = "white") +
        theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
        scale_x_discrete(limits = sorted_pca$School) +
        xlab(NULL) +
        ylab("PCA score")

# ggsave("PCA_scores.png", width = 16, height = 8, units = "in", dpi = 300)
```

Now, we can save the data:

```{r}
write.csv(sorted_pca, "dental_schools_ranking_data.csv", row.names = F)
```
